The Adaptive Multi-scale Simulation Infrastructure - Design and Application

### 2016-01-10 doc/articles/amsi_biotissue/paper.tex
working on paper

\usepackage{graphicx}
\usepackage{mathtools}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

@@

\label{amsi_scales}
A configuration file is supplied during command invocation containing the initial configuration of the process sets and their assignment to individual scales, as well as declaring the coupling between scales in the simulation. The deceleration of a scale and the initialization of the process set associated with the scale are considered a single atomic operation in the current implementation.

@@

Deceleration of a coupling relation between two scales is required if any communication between the scales is to occur during the simulation (through the amsi libraries), as the coupling deceleration during simulation initialization is used to instantiate data structures specific to scale-coupling operations. This limits memory overhead by not allocating space for unneeded coupling operations.

Scales are declared to be related if some quantities of interest (such as physical tensor field values) are transmitted between them for scale-coupling. This coupling relationship is assumed to be asymmetric, thus if scale A is related to scale B, scale B is not implicitly related to scale A, though this may be separately declared. For brevity this may be written as A~B for A is coupled to B.

@@

All communication and planning operations in the amsi libraries use an implicit local numbering for data distribution terms, the index associated with the term in a local serialized buffer. At present an application developer must implement up-/down-scaling operations required for their particular multi-scale use case, hooks will be provided in a future release to automate application of such operations during scale-coupling.

\subsection{Soft Tissue Simulation}\label{biotissue}
The Biotissue code is a multi-scale simulation of soft organic tissues, consisting of an engineering-scale simulation using the finite element method, and a micro-scale quasistatics force-balance simulation. The Cauchy Momentum Balance Equation for a body in static equilibrium is used as the governing equation for the macro-scale simulation. 
%
\begin{align}
\nabla \cdot \pmb{\sigma} =& \; 0 \text{ in } \Omega \label{momentum_balance} \\
\pmb{\sigma} \cdot \pmb{n} =& \; \pmb{t} \text{ on } \partial \Omega \nonumber
\end{align}
%
Specific quantities needed to compute the elemental tangent stiffness matrices and force vectors are instead supplied by the micro-scale force-balance simulations, which occur at each numerical integration point in the engineering-scale mesh \ref{biotissue_hierarchy}.

\begin{figure}
  \begin{center}
    \includegraphics[height=2in]{biotissue_visual_hierarchy.eps}
  \end{center}
  \caption{\small biotissue multi-scale problem}
  \label{biotissue_hierarchy}
\end{figure}

@@

Thus the simulation as a whole has two Newton-Raphson convergence operations occurring. During each macro-scale Newton iteration every micro-scale RVE must undergo a full Newton-Raphson convergence process.

\section{Results and Discussion}\label{results}
The biotissue simulation was run on the AMOS BG/Q machine housed at the CCI facilities. Both strong and weak scaling studies were conducted up to an eighth of the total machine core count, or 16384 cores. The ratio of macro-scale to micro-scale assigned processes was held near 1:128, though the total core count was kept as an integer power of 2 in order for the SLURM job scheduler to submit the jobs to the compute nodes in a timely manner. The precise calculation used to determine macro and micro ratio is given by 
%
num\_macro = N 
num\_cores = N*128
num\_micro = num\_cores - N
%
thus as the number of cores assigned to the macro-scale computation increases, the ratio approaches 1:128.

@@

\begin{figure}
  \begin{center}
    \includegraphics[height=2in]{ws_phil.eps}
  \end{center}
  \caption{\small weak scaling}
  \label{weak scaling}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[height=2in]{ss_phil.eps}
  \end{center}
  \caption{\small strong scaling}
  \label{strong scaling}
\end{figure}
As the problem size is held constant with respect to the number of individual elements in the engineering scale domain discretization the simulation displays near optimal convergence to the macro-scale solve time alone, which is also reducing in time cost. The limiting factor in the speedup of the overall simulation is the potential speedup of the macro-scale simulation rather than the micro-scale simulation with fiber-only RVEs. As indicated in the weak scaling results this is related primarily to the macro-scale linear solve communication time increasing. At present an incomplete LU factorization is used to solve the linear system formulated at macro-scale each Newton iteraion (using PETSc configured with the superLU\_dist solver \cite{li05superlu} \cite{lishao10superluilu} \cite{lidemmel03superludist} ). Using instead an iterative solution method to reach a desired level of accuracy may result in stronger scaling at macro-scale, but as yet this approach has been unneeded due to the preponderance of time spent computing micro-scale results in use cases.

>>>>

\usepackage{graphicx}
\usepackage{listings}
\usepackage{mathtools}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

@@

\label{amsi_scales}
A configuration file is supplied during command invocation containing the initial configuration of the process sets and their assignment to individual scales, as well as declaring the coupling between scales in the simulation. The declaration of a scale and the initialization of the process set associated with the scale are considered a single atomic operation in the current implementation. A simple configuration file can be seen in listing \ref{amsi config file} which declares the existence of two scales, macro and micro\_fo, and assigns them 64 and 8128 processes respectively (for a total of $2^13$ processes). The file also specifies the scale-coupling relations, with macro coupled to micro and micro coupled to macro. Note that the order of specification is not relevant to the operation of amsi.
%
\begin{lstlisting}
  macro 64
  micro_fo 8128
  #
  macro micro_fo
  micro_fo macro_fo
\end{lstlisting}
\label{amsi config file}
%

@@

Declaration of a coupling relation between two scales is required if any communication between the scales is to occur during the simulation (through the amsi libraries), as the coupling declaration during simulation initialization is used to instantiate data structures specific to scale-coupling operations. This limits memory overhead by not allocating space for unneeded coupling operations.

Scales are declared to be related if some quantities of interest (such as physical tensor field values) are transmitted between them for scale-coupling. This coupling relationship is assumed to be asymmetric, thus if scale A is related to scale B, scale B is not implicitly related to scale A, though this may be separately declared.

@@

All communication and planning operations in the amsi libraries use an implicit local numbering for data distribution terms, the index associated with the term in a local serialized buffer. At present an application developer must implement up-/down-scaling operations required for their particular multi-scale use case, hooks will be provided in a future release to automate application of such operations during scale-coupling. The current implementation includes on blocking versions of the scale-coupling communication operations. Asynchronous versions have been designed but are not currently being implemented due to lack of immediate need for the primary test amsi use case.

\subsection{Soft Tissue Simulation}\label{biotissue}
The Biotissue code is a multi-scale simulation of soft organic tissues, consisting of an engineering-scale simulation using the finite element method, and a micro-scale quasistatics force-balance simulation. The Cauchy Momentum Balance Equation for a body in static equilibrium is used as the governing equation for the macro-scale simulation. 
%
\begin{align}
\nabla \cdot \pmb{\sigma} =& \; 0 \text{ in } \Omega \label{momentum_balance} \nonumber \\
\pmb{\sigma} \cdot \pmb{n} =& \; \pmb{t} \text{ on } \partial \Omega \nonumber
\end{align}
%
Specific quantities needed to compute the elemental tangent stiffness matrices and force vectors are instead supplied by the micro-scale force-balance simulations, which occur at each numerical integration point in the engineering-scale mesh \ref{biotissue_hierarchy}. 
%
\begin{figure}
  \begin{center}
    \includegraphics[height=2in]{biotissue_visual_hierarchy.eps}
  \end{center}
  \caption{\small biotissue multi-scale problem}
  \label{biotissue_hierarchy}
\end{figure}
%

@@

The simulation two nested Newton-Raphson operations occurring; during each macro-scale Newton iteration every micro-scale RVE must undergo a full Newton-Raphson convergence process.

\section{Results and Discussion}\label{results}
The biotissue simulation was run on the AMOS BG/Q machine housed at the CCI facilities. Both strong and weak scaling studies were conducted up to an eighth of the total machine core count, or 16384 cores. The ratio of macro-scale to micro-scale assigned processes was held near 1:128, though the total core count was kept as an integer power of 2 in order for the SLURM job scheduler to submit the jobs to the compute nodes in a timely manner. The precise calculation used to determine macro and micro ratio is given by 
%
\begin{lstlisting}[Language=C]
  num_cores = 128 * num_macro;
  num_micro = num_core - num_macro;
\end{lstliting}
\label{macro micro ratio}
%
thus as the number of cores assigned to the macro-scale computation increases, the ratio approaches 1:128.

@@

%
\begin{figure}
  \begin{center}
    \includegraphics[height=2in]{ws_phil.eps}
  \end{center}
  \caption{\small weak scaling}
  \label{weak scaling}
\end{figure}
%
Keeping the work per process approximately constant (1k elements per macro-scale process, with the macro:micro ratio as specified in listing \ref{macro micro ratio}) it can be seen that the time to compute every RVE in the system stays relatively consistent, while the time to solve the macro-scale linear system increases, causing a modest increase in the overall time per macro-scale Newton iteration.

%
\begin{figure}
  \begin{center}
    \includegraphics[height=2in]{ss_phil.eps}
  \end{center}
  \caption{\small strong scaling}
  \label{strong scaling}
\end{figure}
%
As the problem size is held constant with respect to the number of individual elements in the engineering scale domain discretization the simulation displays near optimal convergence to the macro-scale solve time alone, which is also reducing in time cost. The limiting factor in the speedup of the overall simulation is the potential speedup of the macro-scale simulation rather than the micro-scale simulation with fiber-only RVEs. As indicated in the weak scaling results this is related primarily to the macro-scale linear solve communication time increasing. At present an incomplete LU factorization is used to solve the linear system formulated at macro-scale each Newton iteraion (using PETSc configured with the superLU\_dist solver \cite{li05superlu} \cite{lishao10superluilu} \cite{lidemmel03superludist} ). Using instead an iterative solution method to reach a desired level of accuracy may result in stronger scaling at macro-scale, but as yet this approach has been unneeded due to the preponderance of time spent computing micro-scale results in use cases.
